---
title: "SN_Project_V1"
output: html_notebook
---

```{r}
# read in relevant libraries
library(data.table)
library(igraph)
library(recommenderlab)
library(ggplot2)
```

```{r}
# set random seed
set.seed(23495)
```

The primary rating data was prepped in a separate code file, sampled, and and stored. The cleaned data is directly imported here for convenience.
```{r}
# read in rating data 
rating.dt <- fread("netflix_sampled_data.csv", header=TRUE) # data for 2004, min. 100 user reviews and 100 movie ratings

# get data on movie names
## adjusted movie title names slightly directly in csv file prior to import
movies.dt <- fread("movie_titles_aws.csv", header=FALSE, col.names=c("MovieID", "Title"))
```

***************

#### Basic summary statistics

```{r}
# number of unique movies
cat("Number of unique movies:", length(unique(rating.dt$MovieID)))

# number of users
cat("\nNumber of users who provided ratings:",length(unique(rating.dt$CustomerID)))

# number of total ratings
cat("\nNumber of total ratings:",nrow(rating.dt))
```

```{r}
# average and median user rating
cat("Average rating across data:",mean(rating.dt$Rating))
cat("\nMedian rating across data:",median(rating.dt$Rating))
```

```{r}
# distribution of average rating by user (indicating lack of uniformity)
# color options: #AA2B2B #9D2E2E ##98141D --> pptx dark red
avg.ratings <- rating.dt[, .(AvgRating=mean(Rating)), by=CustomerID]
ggplot(avg.ratings, aes(x=AvgRating)) + 
  geom_histogram(binwidth=0.2, col="gray", fill="#9D2E2E") + 
  labs(x="Average User Rating", y="Number of Users", title="Distribution of Average User Rating") + theme_minimal() +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("Average rating histogram.png", width=7, height=5)  

```

```{r}
# distribution of movie rating by user
# color options: #AA2B2B #9D2E2E
ggplot(rating.dt, aes(x=Rating)) + 
  geom_histogram(binwidth=0.3, col="gray", fill="#9D2E2E") + 
  labs(x="User Rating of a Movie", y="Number of Users", title="Distribution of User Ratings") + 
  theme_minimal() +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("Rating Histogram.png", width=7, height=5)
```

```{r}
# look at number of movies rated by user

# get table with number of movies rated by each user
user.ratings <- rating.dt[, .("NumRated"=.N), by=CustomerID]
rating.dt[,Avg_rating := mean(Rating), by=CustomerID]
rating.dt[,numbers := .N, by=CustomerID]

rating_per_user <- unique(rating.dt[,.(numbers, Avg_rating)])
rating_per_user_2000 <- rating

hist(rating_per_user$Avg_rating)

ggplot(rating_per_user, aes(numbers, Avg_rating)) + 
  geom_smooth(method = "lm", col="#9D2E2E") + 
  labs(title="Average Rating versus Number of Ratings per User", x="Number of Ratings per User", y="Average Rating") + 
  theme_minimal() + coord_cartesian(ylim=c(2.5,4)) +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("Average rating vs number of ratings per user.png", width=7, height=5)

summary(lm(Avg_rating ~ numbers, data = rating_per_user))

# average number of movies rated
cat("Average number of movies rated:",mean(user.ratings$NumRated))

# median number of movies rated
cat("\nMedian number of movies rated:",median(user.ratings$NumRated))

# max number of movvies rated
cat("\nMax number of movies rated:",max(user.ratings$NumRated))
```
A number of users had very high movie ratings, e.g. in the 4000s that would imply 10+ movies seen per day on average. This may be due to multiple individuals sharing an account, or due to the use of on-site surveys to get ratings of movies a user saw in the past.

```{r}
# get Top10 movies with highest number of ratings
movies.info <- rating.dt[, .("NumberofRatings"=.N, "AvgRating"=mean(Rating)), by=MovieID]
movies.info <- merge(movies.info, movies.dt, by="MovieID")
```

```{r}
# get Top10 movies with highest number of ratings
print("Movies with highest number of ratings")
head(movies.info[order(-NumberofRatings)]$Title, 10)
```

```{r}
# get Top10 movies with highest average rating
print("Movies with highest average ratings")
head(movies.info[order(-AvgRating)]$Title, 10)
```

```{r}
# Full plot
ggplot(movies.info, aes(x=NumberofRatings, y=AvgRating)) + 
  geom_smooth(se=F, col="#9D2E2E", size=1.1) + 
  labs(title="Average Rating versus Movie Degree Centrality", x="Movie Degree Centrality (in bipartite network)", y="Average Rating") + 
  theme_minimal() + scale_x_continuous(breaks=seq(0,60000,10000)) +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("Average rating vs degree.png", width=7, height=5)

ggplot(movies.info, aes(x=log(NumberofRatings), y=AvgRating)) + 
  geom_smooth(method = "lm", col="#9D2E2E") + geom_point() +
  labs(title="Average Rating versus Movie Degree Centrality (logged)", x="Movie Degree Centrality (logged, in bipartite network)", y="Average Rating") + 
  theme_minimal() + coord_cartesian(ylim=c(2.5,4)) +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
cor(movies.info$NumberofRatings, movies.info$AvgRating)

avg.ratings <- rating.dt[, .(AvgRating=mean(Rating)), by=CustomerID]
ggplot(avg.ratings, aes(x=AvgRating)) + 
  geom_histogram(binwidth=0.2, col="gray", fill="#9D2E2E") + 
  labs(x="Average User Rating", y="Number of Users", title="Distribution of Average User Rating") + theme_minimal() +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

# distribution of unlogged number of ratings
ggplot(movies.info, aes(x=NumberofRatings)) + 
  geom_histogram(binwidth=200, col="grey", fill="#9D2E2E") +
  labs(title="Distribution of Movie Degree Centrality", x="Movie Degree Centrality", y="Number of Ratings") +
  theme_minimal() +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

# distribution of logged number of ratings
ggplot(movies.info, aes(x=log(NumberofRatings))) + 
  geom_histogram(binwidth=0.1, col="grey", fill="#9D2E2E") +
  labs(title="Distribution of Movie Degree Centrality (logged)", x="Movie Degree Centrality (logged)", y="Number of Ratings") +
  theme_minimal() +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.1, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
```

```{r}
# Zoomed In Plot
ggplot(movies.info[NumberofRatings<3000,], aes(x=NumberofRatings, y=AvgRating)) + 
  geom_smooth(method="loess", se=F, col="#9D2E2E", size=1.1) + 
  labs(title="Zoomed in: Average Rating versus Movie Degree Centrality", x="Movie Degree Centrality (in bipartite network)", y="Average Rating") + 
  theme_minimal() + scale_x_continuous(breaks=seq(0,3000,500)) +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("Average rating vs degree_ZOOMED.png", width=7, height=5)

hist(movies.info$NumberofRatings)
hist(log(movies.info$NumberofRatings))
qqnorm(log(movies.info$NumberofRatings))

# scatter plot
ggplot(movies.info[], aes(x=log(NumberofRatings), y=AvgRating)) + 
    geom_smooth(method="lm") +
  labs(title="Zoomed in: Average Rating versus Movie Degree Centrality", x="Movie Degree Centrality (in bipartite network)", y="Average Rating") + 
  theme_minimal() + scale_y_continuous(breaks=seq(2.5,4,0.5), limits =c(2.5, 4)) +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("Average rating vs degree_ZOOMED.png", width=7, height=5)

```

```{r}
# Linear Regression on movies with 1500 or fewer ratings
regress.dt <- movies.info[NumberofRatings <= 1000,] # create subset of movies with 1500 or fewer ratings
setnames(regress.dt, "NumberofRatings", "Degree") # change name to degree
summary(lm(AvgRating~Degree, data=regress.dt)) # regression on subset

summary(lm(AvgRating~log(NumberofRatings), data=movies.info)) # regression on all data
```

```{r}
# distribution of number of movies rated by user, limit axis to 1000+ movies
plot.ratings <- user.ratings[,.(NumRated = ifelse(NumRated>=1000, 1000, NumRated))]
ggplot(plot.ratings, aes(x=NumRated)) + 
    geom_histogram(binwidth=8, col="gray", fill="#9D2E2E") + 
    labs(x="Number of Movies Rated", y="Number of Users", title="Distribution of Number of Movies Rated") + 
    theme_minimal() + scale_x_continuous(breaks=seq(100,1000,100),
                                     labels=c("100","200","300","400","500","600","700","800","900", "1000+")) +
  #scale_y_continuous(breaks=seq(0,60000,10000)) +
    theme(text=element_text(family="Roboto"),
          plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
          axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
          axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("number of movies rated histogram.png", width=7, height=5)
```

#### Building Networks: MOVIE-USER NETWORK

```{r}
# prep for making network
rating.dt[,CustomerID := sub("^", "u", CustomerID )]
rating.dt[, MovieID := as.character(MovieID)]

# make bipartite graph
graph.bp <- graph.data.frame(rating.dt[,1:2], directed=FALSE) # make general undirected graph
V(graph.bp)$type <- V(graph.bp)$name %in% rating.dt$MovieID # specify type to make bipartite
E(graph.bp)$weight <- rating.dt$Rating # add in rating as weight

# look at graph
graph.bp
```

```{r}
customers <- rating.dt[CustomerID %in% sample(unique(CustomerID), 100),]
customers <- customers[MovieID %in% sample(unique(MovieID), 1000),]

graph.customers100 <- graph.data.frame(customers[,1:2], directed=FALSE)
V(graph.customers100)$type <- V(graph.customers100)$name %in% customers$MovieID # specify type to make bipartite
E(graph.customers100)$weight <- customers$Rating # add in rating as weight

graph.customers100
```
```{r}
col <- c("gray85", "#9D2E2E")
shape <- c("circle", "square")

plot(graph.customers100,
     vertex.color = col[as.numeric(V(graph.customers100)$type)+1],
     vertex.shape = shape[as.numeric(V(graph.customers100)$type)+1],  
     layout=layout_as_bipartite(graph.customers100),
     vertex.frame.color="gray60",
     edge.color = "#E5AAAA",
     vertex.label="",
     vertex.label.cex=1, vertex.size=5, edge.width = 0.1,
     asp=0.5)


```


```{r fig.width=20, fig.height=20}
# Visualize graph layout
bp.subplot <- induced_subgraph(graph.bp,v=sample(unlist(V(graph.bp)$name), 7500))
bp.subplot <- subgraph.edges(graph.bp,seq(1,8467727,100))

# define color and shape mappings.
col <- c("gray85", "#9D2E2E")
shape <- c("circle", "square")

plot(bp.subplot,
     vertex.color = col[as.numeric(V(bp.subplot)$type)+1],
     vertex.shape = shape[as.numeric(V(bp.subplot)$type)+1], layout=layout_as_bipartite(bp.subplot, hgap=30),
     vertex.frame.color="gray60",
     edge.color = "#E5AAAA",
     vertex.label="",
     vertex.label.cex=1, vertex.size=5, edge.width = 0.1,
     asp=0.5)

bp.subplot <- subgraph.edges(graph.bp,1:10000)
plot(bp.subplot,
     vertex.color = col[as.numeric(V(bp.subplot)$type)+1],
     vertex.shape = shape[as.numeric(V(bp.subplot)$type)+1], layout=layout_as_bipartite(bp.subplot, hgap=30),
     vertex.frame.color="gray60",
     edge.color = "#E5AAAA",
     vertex.label.color="gray85",
     vertex.label.cex=1.5, vertex.size=5, edge.width = 0.1,
     asp=0.5)

```

#### Building Networks: MOVIE-MOVIE NETWORK

```{r}
# make bipartite graph on movie-movie network
graph.bp2 <- graph.data.frame(rating.dt[,2:1], directed=FALSE) # make general undirected graph
V(graph.bp2)$type <- V(graph.bp2)$name %in% rating.dt$CustomerID # specify type to make bipartite

# look at graph
graph.bp2
```

```{r}
mov.mtx <- as_incidence_matrix(graph.bp2) # get affiliation matrix from chart
mov.sp.mtx <- as(mov.mtx, "sparseMatrix") # encode as sparse matrix
mov.coaffil.mtx <- tcrossprod(mov.sp.mtx) # get co-affiliation matrix to make movie network
```

```{r}
# make movie-movie coaffiliation network
graph.movies <- graph_from_adjacency_matrix(mov.coaffil.mtx, mode="undirected", diag=FALSE, weight=TRUE) # keep diagonals because indicate own rating strength?
```

```{r fig.width=20, fig.height=20}
graph.movies

mov.subplot <- induced_subgraph(graph.movies,v=sample(unlist(V(graph.movies)$name), 1000))
plot.igraph(graph.movies, vertex.label="", vertex.color="gray70", vertex.frame.color="gray30",
            edge.color="#E5AAAA", vertex.size=1, layout=layout_with_kk, edge.width = 0.1)

```

```{r}
# calculate co-affiliation centrality measures
degree.score2 <- degree(graph.movies)
closeness.score2 <- closeness(graph.movies)
eigen.score2 <- eigen_centrality(graph.movies)

movies.ratings.2 <- rating.dt[MovieID %in% V(graph.movies)$name,
                               .("AvgRating"=mean(Rating)), by=MovieID]
movies.performance.2 <- data.table("MovieID"=V(graph.movies)$name, "Degree"=degree.score2,
                                   "Closeness"=closeness.score2)
movies.performance.2 <- data.table("MovieID"=V(graph.movies)$name, "Degree"=degree.score2, 
                                   "Closeness"=closeness.score2, "EigenCentrality"=eigen.score2$vector)
movies.performance.2 <- merge(movies.performance.2, movies.ratings.2, by="MovieID")
```

```{r}
# Movie-Movie Degree Centrality
ggplot(movies.performance.2, aes(x=Degree, y=AvgRating)) +
  geom_smooth(method="loess", se=F, col="#9D2E2E", size=1.1) +
  labs(title="Average Rating versus Movie-Movie Degree Centrality", x="Movie Degree Centrality (in co-affiliation network)", y="Average Rating") +
  theme_minimal() +
  theme(plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("movie-movie degree vs avg rating.png", width=7, height=5)
hist(log(movies.performance.2$Degree),breaks = 300)

# scatter plot
ggplot(movies.performance.2, aes(x=Degree, y=AvgRating)) +
  geom_point(col="#9D2E2E") +
  labs(title="Average Rating versus Movie-Movie Degree Centrality", x="Movie Degree Centrality (in co-affiliation network)", y="Average Rating") +
  theme_minimal() +
  theme(plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

```

```{r}
# Movie-Movie Closeness Centrality
ggplot(movies.performance.2, aes(x=Closeness, y=AvgRating)) +
  geom_smooth(method="loess", se=F, col="#9D2E2E", size=1.1) +
  labs(title="Average Rating versus Movie-Movie Closeness Centrality", x="Movie Closeness Centrality (in co-affiliation network)", y="Average Rating") +
  theme_minimal() +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("movie-movie closeness vs avg rating.png", width=7, height=5)
hist(movies.performance.2$Closeness)

# scatter plot
ggplot(movies.performance.2, aes(x=Closeness, y=AvgRating)) +
  geom_smooth(method = "lm", col="#9D2E2E") +
  labs(title="Average Rating versus Movie-Movie Closeness Centrality", x="Movie Closeness Centrality (in co-affiliation network)", y="Average Rating") +
  theme_minimal() + coord_cartesian(ylim=c(2.5,4)) +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
```

```{r}
# # Movie-Movie Eigen Centrality
ggplot(movies.performance.2, aes(x=EigenCentrality, y=AvgRating)) +
  geom_smooth(method="loess", se=F, col="#9D2E2E", size=1.1) +
  labs(title="Average Rating versus Movie-Movie Eigen Centrality", x="Movie Eigen Centrality (in co-affiliation network)", y="Average Rating") +
  theme_minimal() 
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("movie-movie eigen vs avg rating.png", width=7, height=5)

hist(movies.performance.2$EigenCentrality, breaks=30)
hist(log(movies.performance.2$EigenCentrality), breaks=30)

ggplot(movies.performance.2, aes(x=log(EigenCentrality), y=AvgRating)) +
  geom_smooth(method = "lm", col = "#9D2E2E") +
  labs(title="Average Rating versus Movie-Movie Eigen Centrality (logged)", x="Movie Eigen Centrality (logged, in co-affiliation network)", y="Average Rating") +
  theme_minimal() + coord_cartesian(ylim=c(2.5,4)) +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

# distribution of unlogged Movie Eigen Centrality
ggplot(movies.performance.2, aes(x=EigenCentrality)) + 
  geom_histogram(binwidth=0.05, col="grey", fill="#9D2E2E") +
  labs(title="Distribution of Movie Eigen Centrality", x="Movie Eigen Centrality", y="Number of Ratings") +
  theme_minimal() +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

# distribution of logged Movie Eigen Centrality
ggplot(movies.performance.2, aes(x=log(EigenCentrality))) + 
  geom_histogram(binwidth=0.3, col="grey", fill="#9D2E2E") +
  labs(title="Distribution of Movie Eigen Centrality (logged)", x="Movie Eigen Centrality (logged)", y="Number of Ratings") +
  theme_minimal() +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.1, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
```

```{r}
# how to intepret
summary(lm(AvgRating ~ Degree, data = movies.performance.2)) # degree
summary(lm(AvgRating ~ Closeness, data = movies.performance.2)) # closeness
summary(lm(AvgRating ~ log(EigenCentrality), data = movies.performance.2)) # EigenCentrality
```

#### Building Networks: USER-USER NETWORK

```{r}
# limit user-user network to users with at least 200 ratings
users.sample <- copy(rating.dt)
users.sample <- users.sample[, NumRatings := .N, by=CustomerID]
users.sample <- users.sample[NumRatings >= 200, ]
users.sample <- users.sample[NumRatings < 4000, ]

# create bipartite graph with less data
graph.bp3 <- graph.data.frame(users.sample[,1:2], directed=FALSE) # make general undirected graph
V(graph.bp3)$type <- V(graph.bp3)$name %in% users.sample$MovieID # specify type to make bipartite
E(graph.bp3)$weight <- users.sample$Rating # add in rating as weight

# get coaffiliation matrix from bipartite graph
users.mtx <- as_incidence_matrix(graph.bp3) # get affiliation matrix from chart
users.sp.mtx <- as(users.mtx, "sparseMatrix") # encode as sparse matrix
users.coaffil.mtx <- tcrossprod(users.sp.mtx) # get co-affiliation matrix to make movie network
```

```{r}
# make user-user coaffiliation network
graph.users <- graph_from_adjacency_matrix(users.coaffil.mtx, mode="undirected", diag=FALSE, weight=TRUE) # keep diagonals because indicate own rating strength?
```

```{r fig.width=20, fig.height=20}
#graph.users
users.subplot <- induced_subgraph(graph.users,v=sample(unlist(V(graph.users)$name), 2000))
plot.igraph(users.subplot, vertex.color="gray70", vertex.label="", vertex.frame.color="gray30",
            edge.color="#E5AAAA", vertex.size=2, layout=layout_with_kk, edge.width=0.1)
```

```{r}
# calculate co-affiliation centrality measures
degree.score3 <- degree(graph.users)
closeness.score3 <- closeness(graph.users)
eigen.score3 <- eigen_centrality(graph.users)

movies.ratings.3 <- rating.dt[CustomerID %in% V(graph.users)$name,
                               .("AvgRating"=mean(Rating)), by=CustomerID]
movies.performance.3 <- data.table("CustomerID"=V(graph.users)$name, "Degree"=degree.score3,
                                   "Closeness"=closeness.score3)
movies.performance.3 <- data.table("CustomerID"=V(graph.users)$name, "Degree"=degree.score3,
                                   "Closeness"=closeness.score3, "EigenCentrality"=eigen.score3$vector)
movies.performance.3 <- merge(movies.performance.3, movies.ratings.3, by="CustomerID")
```

```{r}
# User-User Degree Centrality
ggplot(movies.performance.3, aes(x=Degree, y=AvgRating)) +
  geom_smooth(method="loess", se=F, col="#9D2E2E", size=1.1) +
  labs(title="Average Rating versus User-User Degree Centrality", x="User Degree Centrality (in co-affiliation network)", y="Average Rating") +
  theme_minimal() +
  theme(plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("user-user degree vs avg rating.png", width=7, height=5)

hist(movies.performance.3$Degree)

ggplot(movies.performance.3, aes(x=Degree, y=AvgRating)) +
  geom_point(col="#9D2E2E") +
  labs(title="Average Rating versus User-User Degree Centrality", x="User Degree Centrality (in co-affiliation network)", y="Average Rating") +
  theme_minimal() +
  theme(plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
```

```{r}
# User-User Closeness Centrality
ggplot(movies.performance.3, aes(x=Closeness, y=AvgRating)) +
  geom_smooth(method="lm", col="#9D2E2E") +
  labs(title="Average Rating versus User-User Closeness Centrality", x="User Closeness Centrality (in co-affiliation network)", y="Average Rating") +
  theme_minimal() + coord_cartesian(ylim=c(2.5,4)) +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("user-user closeness vs avg rating.png", width=7, height=5)

hist(movies.performance.3$Closeness)

ggplot(movies.performance.3, aes(x=Closeness, y=AvgRating)) +
  geom_smooth(method = "lm", col = "#9D2E2E") +
  labs(title="Average Rating versus User-User Closeness Centrality", x="User Closeness Centrality (in co-affiliation network)", y="Average Rating") + geom_smooth(method = "lm") +
  theme_minimal() + coord_cartesian(ylim=c(2.5,4)) +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

ggsave("user-user closeness vs avg rating.png", width=7, height=5)

```

```{r}
# # Movie-Movie Eigen Centrality
ggplot(movies.performance.3, aes(x=EigenCentrality, y=AvgRating)) +
  geom_smooth(method="loess", se=F, col="#9D2E2E", size=1.1) +
  labs(title="Average Rating versus User-User Eigen Centrality", x="User Eigen Centrality (in co-affiliation network)", y="Average Rating") +
  theme_minimal() +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("user-user eigen vs avg rating.png", width=7, height=5)

hist(log(movies.performance.3$EigenCentrality))

ggplot(movies.performance.3, aes(x=log(EigenCentrality), y=AvgRating)) +
  geom_point(col="#9D2E2E") + geom_smooth(method = "lm") +
  labs(title="Average Rating versus User-User Eigen Centrality", x="User Eigen Centrality (in co-affiliation network)", y="Average Rating") +
  theme_minimal() +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
ggsave("user-user eigen vs avg rating.png", width=7, height=5)

ggplot(movies.performance.3, aes(x=log(EigenCentrality), y=AvgRating)) + 
  geom_smooth(method = "lm", col = "#9D2E2E") +
  labs(title="Average Rating versus User-User Eigen Centrality (logged)", x="User Eigen Centrality (logged, in co-affiliation network)", y="Average Rating") +
  theme_minimal() + coord_cartesian(ylim=c(2.5,4)) +
  theme(text=element_text(family="Roboto"),
        plot.title=element_text(size=14, hjust=0.5, margin = margin(t = 5, r = 0, b = 8, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 12, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
```

```{r}
# how to intepret
summary(lm(AvgRating ~ Degree, data = movies.performance.3)) # degree
summary(lm(AvgRating ~ Closeness, data = movies.performance.3)) # closeness
summary(lm(AvgRating ~ log(EigenCentrality), data = movies.performance.3)) # EigenCentrality
```

#### RECOMMENDER LAB

```{r}
# get item-movie matrix from directed graph
input.mtx <- as_incidence_matrix(graph.bp, attr="weight", sparse=TRUE)

# store as recommender lab matrix #and normalize
input.mtx <- as(input.mtx, "realRatingMatrix")
```

```{r}
# inspect the rating distributions
hist(getRatings(input.mtx))
hist(getRatings(normalize(input.mtx)), breaks=100)
hist(getRatings(normalize(input.mtx, method="Z-score")), breaks=100)
hist(colMeans(input.mtx), breaks=20)
```

# Test example to produce sample recommendations (updated to give random sample)
```{r}
# create a recommender on UBCF
rec.model = Recommender(input.mtx, method = "UBCF")

# get a random user name
test.user <- sample(rating.dt$CustomerID, 1)

# get the top 10 movies recommended for user XX
rowCounts(input.mtx[test.user])

# what he rated high
rated.high <- rating.dt[CustomerID==test.user & Rating > 3,]

rec.test.user = predict(rec.model, input.mtx[test.user,], n=10)
rec.compare <- as.numeric(unlist(as(rec.test.user, "list")))

# movies he rated high
high <- movies.dt[MovieID %in% rated.high$MovieID, .(MovieID,Title)]

# movies recommended
recommend <- movies.dt[MovieID %in% rec.compare, .(MovieID,Title)]

high
recommend
```

# Evaluation
```{r}
# evaluate different methods
eval = evaluationScheme(input.mtx, method="split", train=0.75, given = 20, goodRating = 4)
eval

# algorithms (perform normalization automatically)
algorithms <- list(
  "random items" = list(name="RANDOM", param=NULL),
  "popular items" = list(name="POPULAR", param=NULL),
  "user-based CF" = list(name="UBCF", param=list(nn=50)),
  "item-based CF" = list(name="IBCF", param=list(k=50)),
  "SVD approximation" = list(name="SVD", param=list(k = 50)))
```

```{r}
# evaluate top-N recommendations
results1 <- evaluate(eval, algorithms, type = "topNList", n=c(1, 5, 10, 20, 50))
```
```{r}
# ROC Curve
plot(main="Comparison of ROC curves for 5 recommender methods", results1, legend="topleft", col=c("#231F20", "#00680A", "#140152", "#2274A5", "#9D2E2E"), cex=0.8, lwd=1.2, annotate=c(5))
```

```{r}
# precision-recall curve
plot(y="prec/rec", main="Comparison of ROC curves for 5 recommender methods", results1, legend="bottomright", col=c("#231F20", "#00680A", "#140152", "#2274A5", "#9D2E2E"), cex=0.8, lwd=1.2, annotate=c(5))
```


```{r}
# evaluate ratings prediction
results2 <- evaluate(eval, algorithms, type = "ratings")

# MSE / MAE plot for rating
plot(results2, ylim = c(0,3), main="Comparison of RMSE, MSE, and MAE for 5 recommender methods")
```
```{r}
# MSE / MAE plot for rating
plot(results2, ylim = c(0,3), main="Comparison of RMSE, MSE, and MAE for 5 recommender methods",
col=c("#AFAFAF", "#67936B", "#314E89", "#7AA4BC", "#9E4545"))
```



